pdp: An R Package for Constructing Partial Dependence Plots

### 2016-12-16 manuscript/reviews/response.Rmd

#### Comment
Consider adding a `slowtests` directory (below `pdp/inst/` or elsewhere) that has tests for all the models supported by the package (Table 1 in the paper) to easily check back compatibility when packages change.  (For justification for such tests see for example my comments on `gbm` back compatibility above.)  These tests will probably be quite slow, but that won't be an issue because they won't be invoked by CRAN check; the slow tests would be just for the maintainer of the `pdp` package to run manually.

#### Response
This is a good suggestion and is currently under development. In fact, the `gbm` maintainers have already contacted me about a breaking change caused by the next planned release of their package.

## ----------Reviewer 2----------

#### Comment 
The Introduction immediately introduces the PDP. You must first motivate why it is important. In a linear model, they are not important. In a linear model with simple interactions, likewise. They first become important if you have non-parametric machine-learning techniques. Further, they are not the only idea extant when trying to explain a predictor's impact on the response. This paper would benefit if you can give a bit more of a lit review here.
#### Response
I agree that PDPs are not as useful in "simple" linear models (e.g., few variables, only two-way interactions, etc.), but certainly there will be cases where PDPs will be extremely useful (e.g., many terms, complex interactions, etc.). In fact, linear models are still common in machine learning/computer science/data science (they just typically contain many terms). I also agree that PDPs are not the only, or sometimes best, approach to understanding the impact of a subset of the predictor's on the response. Brief descriptions and references are given for many alternatives: marginal model plots (`plotmo`), individual conditional expectation plots (`ICEbox`)---which are described as an alternatve to PDPs when strong interactions are suspected, effect plots (`effects`), and a few others available in the `car` package. Unless the editor disagrees, I don't think it is necessary to go into more detail than what is already provided at the end of the introduction. Nonetheless, I added a bit more of a motivational introduction that links to the examples using the Boston housing data.

@@

### Important Clarifications Required:

#### Comment
Top of p2. The `ICEbox` package allows for general construction and visualization of PDP's. However, it is limited as you only get one dimensional representations. The selling point of your package should focus on the flexibility of graphics and the illustration of non-extrapoloation zones.
#### Response
ICEbox does not actually construct PDPs. Only graphics trying to do the same thing: visualize the relationship between a predictor and the response. I think flexible graphics and non-extrapoloation zones are only half of the selling point here. I think that being able to run `partial` in parallel is huge, and also that this packge works with many types of machine learning models---not just tree-based models. For example, as described in the manuscript, `pdp` can produce single or multi-predictor PDPs for conditional random forests and XGBoost models (among many others), and in parallel. I expanded the last example to describe the interface with `caret`, which I think opens the door for wonderful exploratory data analysis pipe lines (e.g., choose model --> plot importance --> plot relationships, etc.). Most, if any, of these other software packages do not describe or support these important components (e.g., parallel execution), not even ICEbox.

#### Comment
All `pdp` figures throughout the paper: Can you make explicit in the caption what code produces the left and right subfigures? I would use the subfigure package in latex and then put a comment in the R code e.g. # Fig 3b.
#### Response
Rather than using subfigures, which I think would look messy with all the code, I made quite an attempt to comment which pieces of code produce which figure. Also, in response to another reviewer's comments, the code which produces figures containing multiple plots is better described and demonstrated. See, for example, the section called Multi-predictor PDPs.

@@

### Edits:

#### Comment
p1 abstract: strict assumptions imposed by traditional statistical models
#### Response
Added some clarification here; in particular, normality, constant variance, and linearity.

@@

#### Comment
Fig 7: as beautiful as this is, you should label the sidebar as "logit" or "centered logit" instead of "y".
#### Response
Fixed. Color key labeled "Average logit".

### Suggestions for the software

#### Comment
I understand that you like the feature of having a one-shot function partial to both generate and plot the PDP. I realize that the default is `FALS`E. However, I still think it's a disservice to the user as constructing PDP's is computationally intense and if they get the plot wrong or it's not to their liking, they will have to regenerate the whole thing again. So, it is my opinion that you should take it out and force two steps: generation and illustration.
#### Response
While I am not suggesting that the `partial` function needs to mimic the behavior of the PDP fucntions in `randomForest` and `gbm`, I do find it useful to be able to plot the results without forcing the user to store them. However, I agree with your concern which is why `partial` has `plot = FALSE` as the default. Furthermore, I have made it possible to retrieve the raw PDP data even when \code{plot = TRUE}. This is because the "trellis" object returned by `partial` \ `plotPartial` now contains an additional `partial.data` attribute. This is described in the package documentation. For example,
```{r, eval=TRUE}
library(pdp)
data(trees)
mod <- lm(Volume ~ ., data = trees)
partial(mod, pred.var = "Girth", plot = TRUE)
pdp.Girth <- lattice::trellis.last.object()
pd.Girth <- attr(pdp.Girth, "partial.data")
head(pd.Girth)
```

#### Comment
The package could be more powerful if you allowed a predict function as a parameter. For instance, you cannot currently use random forests to illustrate classification probabilities.
#### Response
I agree, and I have changed the behavior of `...` to refer to additional parameters to `stats::predict` (e.g., the `n.trees` argument in `predict.gbm`.). However, I'm not sure yet about the scale on which the predictions are plotted for classification problems. For now, I am more comfortable leaving it as the usual logit scale defined for PDPs (mostly so that they are defined the same whatever the number of class outcomes). But I am not opposed to making this (simple) change in the future. I have added this as a feature request: https://github.com/bgreenwell/pdp/issues/21.

### 2016-12-16 manuscript/greenwell.tex
tweaks based on review comments

\citet{harrison-1978-hedonic} were among the first to analyze the well-known Boston housing data. One of their goals was to find a housing value equation using data on median housing values from $n = 506$ census tracts in the suburbs of Boston from the 1970 census; see Table IV of \citet{harrison-1978-hedonic} for a description of each variable. The data violate many classical assumptions like linearity, normality, and constant variance. Nonetheless, \citeauthor{harrison-1978-hedonic}---using a combination of transformations, significance testing, and grid searches---were able to find a reasonable fitting model ($R^2 = 0.81$). Part of the payoff for there efforts was an interpretable prediction equation which is reproduced in Equation~\eqref{eqn:boston}.
\begin{equation}
\label{eqn:boston}
\begin{split}
\widehat{\log\left(MV\right)} &= 9.76 + 0.0063 RM^2 + 8.98\times10^{-5} AGE - 0.19\log\left(DIS\right) + 0.096\log\left(RAD\right) \\
  & \quad - 4.20\times10^{-4} TAX - 0.031 PTRATIO + 0.36\left(B - 0.63\right)^2 - 0.37\log\left(LSTAT\right) \\
  & \quad - 0.012 CRIM + 8.03\time10^{-5} ZN + 2.41\times10^{-4}  INDUS + 0.088 CHAS - 0.0064 NOX^2
\end{split}
\end{equation}

Nowadays, supervised learning algorithms can fit the data automatically in seconds---typically with higher accuracy. (We will revisit the Boston housing data in Section~\ref{sec:boston}) The downfall, however, is some loss of interpretation; for example, there is usually no simple prediction formula like in Equation~\eqref{eqn:boston}. Fortunately, these methods can still provide insight into the data, but it is not in the form of simple equations. For example, many modern methods, like boosting and random forests, can assign importance scores to all of the predictors in the training data.

@@

For illustration, we will use the (corrected) Boston housing data which are available from the \CRANpkg{mlbench} package \citep{mlbench-pkg}. These data contain the median value of owner-occupied homes in 506 U.S. census tracts in the Boston area, along with 13 independent variables such as the per capita crime rate by town. We begin by loading the data and omitting unimportant columns.
\begin{example}
data(BostonHousing2, package = "mlbench")  # mlbench must be installed!
boston <- BostonHousing2[, -c(1, 2, 5)]
\end{example}
Next, we fit a random forest to the entire data set with default tuning parameters and 500 trees:
\begin{example}
library(randomForest)
set.seed(101)  # for reproducibility
boston.rf <- randomForest(cmedv ~ ., data = boston, importance = TRUE)
varImpPlot(boston.rf)  # Figure 1
\end{example}
The model fit is reasonable, with an \dfn{out-of-bag} (pseudo) $R^2$ of 0.89. The variable importance scores are displayed in Figure~\ref{fig:plotmo_vs_partial}. Both plots indicate that the percentage of lower status of the population (\code{lstat}) and the average number of rooms per dwelling (\code{rm}) are highly associated with the median value of owner-occupied homes (\code{cmedv}). The question then arises, "What is the nature of these associations?" To help answer this, we can look at the partial dependence of \code{cmedv} on \code{lstat} and \code{rm}, both individually and together.

>>>>

\citet{harrison-1978-hedonic} were among the first to analyze the well-known Boston housing data. One of their goals was to find a housing value equation using data on median housing values from $n = 506$ census tracts in the suburbs of Boston from the 1970 census; see Table IV of \citet{harrison-1978-hedonic} for a description of each variable. The data violate many classical assumptions like linearity, normality, and constant variance. Nonetheless, \citeauthor{harrison-1978-hedonic}---using a combination of transformations, significance testing, and grid searches---were able to find a reasonable fitting model ($R^2 = 0.81$). Part of the payoff for there efforts was an interpretable prediction equation which is reproduced in Equation~\eqref{eqn:boston}.
\begin{equation}
\label{eqn:boston}
\begin{split}
\widehat{\log\left(MV\right)} &= 9.76 + 0.0063 RM^2 + 8.98\times10^{-5} AGE - 0.19\log\left(DIS\right) + 0.096\log\left(RAD\right) \\
  & \quad - 4.20\times10^{-4} TAX - 0.031 PTRATIO + 0.36\left(B - 0.63\right)^2 - 0.37\log\left(LSTAT\right) \\
  & \quad - 0.012 CRIM + 8.03\times10^{-5} ZN + 2.41\times10^{-4} INDUS + 0.088 CHAS - 0.0064 NOX^2
\end{split}
\end{equation}

Nowadays, supervised learning algorithms can fit the data automatically in seconds---typically with higher accuracy. (We will revisit the Boston housing data in Section~\ref{sec:boston}.) The downfall, however, is some loss of interpretation; for example, there is usually no simple prediction formula like in Equation~\eqref{eqn:boston}. Fortunately, these methods can still provide insight into the data, but it is not in the form of simple equations. For example, many modern methods, like boosting and random forests, can assign importance scores to all of the predictors in the training data.

@@

For illustration, we will use the (corrected) Boston housing data which are available from the \CRANpkg{mlbench} package \citep{mlbench-pkg}. These data contain the median value of owner-occupied homes in 506 U.S. census tracts in the Boston area, along with 13 independent variables such as the per capita crime rate by town. These are the same data analyzed in \citet{harrison-1978-hedonic}, but with a few corrections and additional spatial variables. We begin by loading the data and omitting unimportant columns.
\begin{example}
data(BostonHousing2, package = "mlbench")  # mlbench must be installed!
boston <- BostonHousing2[, -c(1, 2, 5)]
\end{example}
Next, we fit a random forest to the entire data set with default tuning parameters and 500 trees:
\begin{example}
library(randomForest)
set.seed(101)  # for reproducibility
boston.rf <- randomForest(cmedv ~ ., data = boston, importance = TRUE)
varImpPlot(boston.rf)  # Figure 1
\end{example}
The model fit is reasonable, with an \dfn{out-of-bag} (pseudo) $R^2$ of 0.89. The variable importance scores are displayed in Figure~\ref{fig:plotmo_vs_partial}. Both plots indicate that the percentage of lower status of the population (\code{lstat}) and the average number of rooms per dwelling (\code{rm}) are highly associated with the median value of owner-occupied homes (\code{cmedv}). The question then arises, "What is the nature of these associations?" To help answer this, we can look at the partial dependence of \code{cmedv} on \code{lstat} and \code{rm}, both individually and together.
